---
title: "Getting started with Negative Binomial Regression Modeling"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

This rmarkdown notebook follows the same analysis as in the following link, but in a more tidyverse-fashion:

https://data.library.virginia.edu/getting-started-with-negative-binomial-regression-modeling/
```{r setup}
library(tidyverse)
library(MASS)
```

The following data is from "Categorial Data Analysis", by Alan Agresti (2002) in Table 13.6 in section 13.4.3. The data are from a survey of 1308 people in which they were asked how many homicide victims they know.
```{r data}
black <- c(119,16,12,7,3,2,0)
white <- c(1070,60,14,4,0,0,1)
resp <- c(rep(0:6,times=black), rep(0:6,times=white))
race <- factor(c(rep("black", sum(black)), rep("white", sum(white))), levels = c("white","black"))
victim <- tibble(resp, race)
```

First, notice most respondents are white.
```{r}
victim %>% group_by(race) %>% summarise(count=n())
```
Blacks have a higher mean count than whites.
```{r}
victim %>% group_by(race) %>% summarise(mean=mean(resp))
```
For each race, the sample variance is roughly double the mean; it appears we have overdispersion.
```{r}
victim %>% group_by(race) %>% summarise(variance=var(resp))
```

Take one look at the distribution of counts by race.
```{r}
victim %>% group_by(resp,race) %>% summarise(count = n()) %>% spread(key=race, value=count)
```

Try a Poisson regression.
```{r}
pGLM <- glm(resp ~ race, data=victim, family = poisson)
summary(pGLM)
```

The coefficient 1.73 is the one and only dichotomous predictor. It is the difference in log expected counts.
```{r}
coef(pGLM)
victim %>% group_by(race) %>% summarise(log_mean=log(mean(resp)))
```

We expect to get the sample means when we make a prediction with this model and exponentiate the results.
```{r}
poismeans <- predict(pGLM, newdata = tibble(race=c("white","black"))) %>% exp()
poismeans
```

This says the count of known victims for whites is distributed as Poisson with mean=variance=0.09, and for blacks is distributed as Poisson with mean=variance=0.52. This is not pleasing as we knew a prior that the variance is about twice as much as the mean.

Let's see how bad this is.

```{r}
poisW <- dpois(0:6,lambda = poismeans[1]) * sum(victim$race=="white") 
poisB <- dpois(0:6,lambda = poismeans[2]) * sum(victim$race=="black")
poisTidy <- tibble(
  resp=rep(0:6,4),
  obs=factor(rep(c(rep("observed",7),rep("fitted",7)),2)),
  race=factor(c(rep("white",14),rep("black",14))),
  count=c(white,poisW,black,poisB)) %>%
  mutate(sqrtcount=sqrt(count))
poisTidy %>% ggplot(aes(x=resp,y=sqrtcount)) +
  geom_col(data=subset(poisTidy, obs=='observed'), aes(fill=race), position = "dodge", alpha=0.4) +
  geom_point(data=subset(poisTidy, obs=='fitted'), aes(colour=race)) +
  geom_line(data=subset(poisTidy, obs=='fitted'), aes(colour=race))
```

Let's run a negative binomial model instead. As expected, we get the same coefficients as before.
```{r}
nbGLM <- glm.nb(resp ~ race, data=victim)
summary(nbGLM)
```

Notice, however, that the standard error for the race coefficient is larger (0.2385 instead of 0.14657).

Also notice that the dispersion parameter is 0.2023. We can use it to get the estimated variances for the counts.
```{r}
nbinommeans <- exp(predict(nbGLM, newdata = data.frame(race = c("white","black"))))
#TODO what is a dispersion parameter?
nbinommeans + nbinommeans^2 * (1/nbGLM$theta)
```


```{r}
nbinomW <- dnbinom(x = 0:6, size = nbGLM$theta, mu = exp(coef(nbGLM)[1]))* sum(victim$race=="white") 
nbinomB <- dnbinom(x = 0:6, size = nbGLM$theta, mu = exp(sum(coef(nbGLM))))* sum(victim$race=="black") 
nbinomTidy <- tibble(
  resp=rep(0:6,4),
  obs=factor(rep(c(rep("observed",7),rep("fitted",7)),2)),
  race=factor(c(rep("white",14),rep("black",14))),
  count=c(white,nbinomW,black,nbinomB)) %>%
  mutate(sqrtcount=sqrt(count))
nbinomTidy %>% ggplot(aes(x=resp,y=sqrtcount)) +
  geom_col(data=subset(nbinomTidy, obs=='observed'), aes(fill=race), position = "dodge", alpha=0.4) +
  geom_point(data=subset(nbinomTidy, obs=='fitted'), aes(colour=race)) +
  geom_line(data=subset(nbinomTidy, obs=='fitted'), aes(colour=race))
```